Bootstrap Learning via Modular Concept Discovery

### 2013-04-23 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Minor edits, and fixed one bibtex record.

This example illustrates the central ``bootstrapping'' challenge for
attempts to create intelligent agents without building in vast amounts
of pre-speciﬁed domain knowledge.  Learning from experience is
typically most effective in a system that already knows most of what
it needs to know, because mistakes are most informative.  In a system
that knows almost nothing, where most problem solving attempts
entirely miss their mark, failures are unlikely to shed light on will
work and what will not.  Uninformative solutions vastly outnumber the
rest, and an uninformed agent has no way to know which solutions will
be informative and which will not.  If it is harder to learn the less
an agent knows, how can a learner who knows almost nothing about a domain
ever get off the ground?

Here we propose an answer to this challenge, based on an approach to
discovering new and increasingly sophisticated concepts that build
modularly on previously learned, simpler concepts.  Like many problems
in AI, the bootstrapping challenge and its solution can be cast in
terms of search.  In our circuit-building scenario, the goal is to ﬁnd
circuit wiring procedures that maximize certain criteria.  Search
algorithm typically aim to exploit local or topological structure in
the search space -- here the space of all circuit-wiring procedures.
For example, local search algorithms move from one candidate solution
to a neighboring one. Heuristic search techniques take advantage of
topological structure to prune off large parts of the solution space.
This structure derives from domain-speciﬁc knowledge.  In local
search, domain knowledge provides the similarity metric between
solutions; in heuristic search it provides the heuristic.  Here we ask
how to bootstrap domain-speciﬁc knowledge by extracting search
structure from the space of possible solutions, by attempting to solve
multiple related problems in a given domain at once.

@@

\begin{figure}[t]
\includegraphics{./figures/tablealonedoc.pdf}
\caption{Some learned primitives from the circuit based Boolean
  function learning experiment. We highlight compression using $NOT$,
  $E_1$ and $E_2$. \,(a)The three elementary gates used to generate
  the task set. (b) Two higher-order primitives in the induced
  grammar. Note how $E_1$ can be used to define FALSE by taking the
  NOT function as an argument. (c) Additional common Boolean functions
  found in the learned grammar.  \label{table:booleanCircuits}}
\end{figure}

@@

We constructed two task sets. The first of these was constructed
explicitly to contain familiar modular structure. In this set of
tasks, we sampled $1000$ Boolean circuits using AND, OR, and NOT
gates. To accomplish this, we first sampled either 1, 2 or 3 inputs,
then between 1 and 5 gates, randomly wiring the inputs of each new
gate to the output of one of the existing gates in the circuit. We
continued this sampling procedure until we had 1000 ``connected''
circuits, i.e., circuits all of whose outputs are wired to an input
except for the last one (the output gate). This resulted in $1000$
tasks consisting of $82$ unique Boolean functions, with a distribution
of tasks as shown in Figure~\ref{fig:booldistr}(a). In
Figure~\ref{fig:boolLearningCurvesCircuits}, we present learning
curves for frontier sizes 100, 500 and 10000, using the same procedure
as in Section~\ref{sec:symreg}. In Figure~\ref{fig:booldistr}, we show
that the distribution over Boolean functions enumerated from the
grammar over expressions is much more similar to the true function
distribution after learning than before.

>>>>

This example illustrates the central ``bootstrapping'' challenge for
attempts to create intelligent agents without building in vast amounts
of pre-speciﬁed domain knowledge.  Learning from experience is
typically most effective in a system that already knows much of what
it needs to know, because mistakes are most informative.  In a system
that knows almost nothing, where most problem solving attempts
entirely miss their mark, failures are unlikely to shed light on what
will work and what will not.  Uninformative solutions vastly outnumber
the rest, and an uninformed agent has no way to know which solutions
will be informative and which will not.  If it is harder to learn the
less an agent knows, how can a learner who knows almost nothing about
a domain ever get off the ground?

Here we propose an answer to this challenge, based on an approach to
discovering new and increasingly sophisticated concepts that build
modularly on previously learned, simpler concepts.  Like many problems
in AI, the bootstrapping challenge and its solution can be cast in
terms of search.  In our circuit-building scenario, the goal is to ﬁnd
circuit wiring procedures that maximize certain criteria.  Search
algorithm typically aim to exploit local or topological structure in
the search space -- here the space of all circuit-wiring procedures.
For example, local search algorithms move from one candidate solution
to a neighboring one. Heuristic search techniques take advantage of
topological structure to prune off large parts of the search space.
This structure derives from domain-speciﬁc knowledge.  In local
search, domain knowledge provides the similarity metric between
solutions; in heuristic search it provides the heuristic.  Here we ask
how to bootstrap domain-speciﬁc knowledge by extracting search
structure from the space of possible solutions, by attempting to solve
multiple related problems in a given domain at once.

@@

\begin{figure}[t]
\includegraphics{./figures/tablealonedoc.pdf}
\caption{Some learned primitives from the circuit based Boolean
  function learning experiment. We highlight compression using $NOT$,
  $E_1$ and $E_2$. \,(a) The three elementary gates used to generate
  the task set. (b) Two higher-order primitives in the induced
  grammar. Note how $E_1$ can be used to define FALSE by taking the
  NOT function as an argument. (c) Additional common Boolean functions
  found in the learned grammar.  \label{table:booleanCircuits}}
\end{figure}

@@

We constructed two task sets. The first of these was constructed
explicitly to contain familiar modular structure. In this set of
tasks, we sampled $1000$ Boolean circuits using AND, OR, and NOT
gates. To accomplish this, we first sampled either 1, 2 or 3 inputs,
then between 1 and 5 gates, randomly wiring the inputs of each new
gate to the output of one of the existing gates in the circuit. We
continued this sampling procedure until we had 1000 ``connected''
circuits, i.e., circuits all of whose outputs are wired to an input
except for the last one (the output gate). This resulted in $1000$
tasks consisting of $82$ unique Boolean functions, with a distribution
of tasks as shown in Figure~\ref{fig:booldistr}a. In
Figure~\ref{fig:boolLearningCurvesCircuits}, we present learning
curves for frontier sizes 100, 500 and 10000, using the same procedure
as in Section~\ref{sec:symreg}. In Figure~\ref{fig:booldistr}, we show
that the distribution over Boolean functions enumerated from the
grammar over expressions is much more similar to the true function
distribution after learning than before.
